---
title: "Exercise 3"
author: "Alec Becker"
date: "`r Sys.Date()`"
output: html_document
---

Using a dataset of text descriptions for patent drawings, this script aims to identify the total number of descriptions matching two string patterns.
```{r}
library(stringr)
```

First, the dataset is read into R.
```{r}
df <- read.csv("patent_drawing data.csv")
```

The structure of the dataset is examined. There are 8,156 rows and 3 columns, and the descriptions are included in the `text` column.
```{r}
str(df)
```

There are 8,156 unique description identifiers. This is the same as the total number of descriptions, meaning there are no duplicate descriptions.
```{r}
length(unique(df$uuid))
```

There are 1,096 unique patent identifiers, meaning 1,096 patents are represented across these drawing descriptions. On average, each patent has about `r round((8156 / 1096), 1)` drawing descriptions.
```{r}
length(unique(df$patent_id))
```

There are no missing values across the entire dataset.
```{r}
sum(is.na(df))
```

To standardize and clean the text descriptions, they are changed to entirely lower case letters and all leading and trailing blanks are removed.
```{r}
df$text2 <- str_trim(str_to_lower(df$text))
```

`r sum(str_detect(df$text2, pattern = regex(".*(embod).*")))` descriptions total include the stem "embod-".
```{r}
sum(str_detect(df$text2, pattern = regex(".*(embod).*")))
```

`r sum(str_detect(df$text2, pattern = regex(".*(embod).*(invention).*")))` descriptions total include the stem "embod-" and have the word "invention" occur anywhere after the "embod-" stem. 
```{r}
sum(str_detect(df$text2, pattern = regex(".*(embod).*(invention).*")))
```