---
title: "Exercise 2"
author: "Alec Becker"
date: "`r Sys.Date()`"
output: html_document
---

This file imports the `get_mod_hamming()` function I wrote and passes three test cases to ensure the function is working properly. This function implements a modified version of the [Hamming distance algorithm](https://en.wikipedia.org/wiki/Hamming_distance).

My function returns both the raw total Hamming distance as `dist` and a distance score that is normalized (`dist_norm`) by the number of characters in the strings being compared (i.e., total Hamming distance / number of characters).

The function is imported below.

```{r}
source("hamming_dist.r")
```

The first test case is passed below. The total distance score is `r get_mod_hamming("data Science", "Data Sciency")$dist` and the normalized distance score is `r get_mod_hamming("data Science", "Data Sciency")$dist_norm`.

```{r}
get_mod_hamming("data Science", "Data Sciency")
```

The second test case is passed below. The total distance score is `r get_mod_hamming("organizing", "orGanising")$dist` and the normalized distance score is `r get_mod_hamming("organizing", "orGanising")$dist_norm`.
```{r}
get_mod_hamming("organizing", "orGanising")
```

The third test case is passed below. The total distance score is `r get_mod_hamming("AGPRklafsdyweIllIIgEnXuTggzF", "AgpRkliFZdiweIllIIgENXUTygSF")$dist` and the normalized distance score is `r get_mod_hamming("AGPRklafsdyweIllIIgEnXuTggzF", "AgpRkliFZdiweIllIIgENXUTygSF")$dist_norm`.

```{r}
get_mod_hamming("AGPRklafsdyweIllIIgEnXuTggzF", "AgpRkliFZdiweIllIIgENXUTygSF")
```

The standard Hamming distance algorithm could be useful when comparing each string in a given set of strings with an expected value. For example, a set of strings could be uploaded or imported
from one storage space to another. If it were possible that some of the strings could be corrupted and changed during the transfer process, calculating the Hamming distance between
each individual string and its expected value could be useful. In this scenario, strings with Hamming distances greater than 0 could be flagged and strings with the highest Hamming distances might
warrant the most investigation.